% \vspace{-0.5ex}
\section{Conclusion}
\label{sec:conclusion}
\vspace{-0.5ex}
We present \dataset{},  a large-scale dataset and framework for scalable and high-quality rubric generation.
By decomposing evaluation into hard rules and principles through Contrastive Rubric Generation and applying preference–label consistency filtering, we construct interpretable and discriminative rubric signals that better align with human judgment.
Our rubric-based reward model, \modelname{}, delivers an average 8.4\% improvement across benchmarks and further boosts policy performance on diverse benchmarks with offline reinforcement learning. 
These results position rubrics-as-rewards as a practical foundation for transparent and generalizable LLM alignment. 
% For future work, we will extend rubric generation to more open-ended tasks and leverage rubrics as intermediate supervision within RLHF pipelines.

\section*{Limitations}
While \dataset{} demonstrates strong empirical gains, several limitations remain.
First, our rubric generation relies on contrastive signals from preference data and inherited model judgments. Although preference-label consistency filtering reduces noise, the resulting rubrics may still reflect biases present in the underlying models and datasets, particularly for subjective or culturally nuanced criteria.
Second, our framework focuses on pairwise comparative evaluation; extending rubric-based rewards to absolute scoring or multi-response ranking scenarios remains an open challenge.
Finally, we primarily evaluate rubric-based rewards in offline preference optimization. How such structured rewards interact with fully online RLHF pipelines, exploration dynamics, and long-horizon policy learning remains an important direction for future work.